---
title: "HRD Biomarkers Prisma Diagram"
subtitle: "[github.com/insilica/bawto...fig1-prisma.R](https://github.com/insilica/bawto/blob/main/figure-scripts/fig1-prisma.R)"
date: "`r Sys.time()`"
output: 
  rmdformats::readthedown:
    self_contained: TRUE
    number_sections: FALSE
    code_folding: hide
---

<figure>
<embed src="http://insilica.co.s3.amazonaws.com/bawto/figures/figure 1.svg" type="image/svg+xml" style="width:100%"/>
<figcaption>A prisma diagram for...</figcaption>
</figure>

The above diagram in Frontiers - ... visualizes screening decisions for a document review of prostate cancer clincial trials:  
1. [sysrev.com/p/63101](https://sysrev.com/u/139/p/63101): screening project for clinicaltrials.gov metadata  
2. [sysrev.com/p/68027](https://sysrev.com/o/2/p/68027): screening project for pubmed abstracts  
3. [sysrev.com/p/70431](https://sysrev.com/p/70431): pubmed data extraction project  

The count data is all stored in a named list:  
```{r results=F,message=F, class.source = 'fold-show'}
prisma <- readRDS(url("http://insilica.co.s3.amazonaws.com/bawto/prisma.rds"))
```

The code to generate these counts is in the main analysis notebook at [github.com/insilica/bawto/.../fig1-prisma.Rmd](https://github.com/insilica/bawto/blob/main/figure-scripts/fig1-prisma.R).
Below the code to generate the figure is given. This code can be easily adapted to other, similar, prisma diagrams

## ClinicalTrials.gov Prisma Diagram:
```{r, class.source = 'fold-show'}
clinicaltrials_prisma <- glue::glue(.open="<",.close=">",
  '
    subgraph cluster_0 {
      label="A. ClinicalTrials.gov BAWTO Screen"
      style=filled;
      fillcolor=lightgrey;
      margin=8
      c_id    [label="Records found   \\non clinicaltrials.gov\\n(n=<prisma$ctgov>)"];
      c_scr   [label="Records screened\\n(n=<prisma$ctgov>)"];
      c_excl  [label="Records excluded\\n(n=<prisma$ctgov_excl>)"];
      c_incl  [label="Records included\\n(n=<prisma$ctgov_incl>)"]
      c_doc   [label="Related Records \\nidentified\\n(n=<prisma$ctgov_ass>)", fillcolor=darkslategray1, style=filled];

      c_id -> c_scr;
      c_scr -> c_excl;
      c_scr -> c_incl;
      c_incl  -> c_doc;
    }
  ')

graph_prisma <- function(text){
  DiagrammeR::grViz(glue::glue(.open="<",.close=">",
  'digraph prisma {
    newrank=true;
    node [shape="box", fontsize = 12, width=2];
    graph [compound=true,splines=ortho, nodesep=0.25, rankdir=TB];
    <text>
  }'))
}
graph_prisma(clinicaltrials_prisma)

```

## Pubmed.gov Prisma
Center and right prisma diagram code:
```{r, class.source = 'fold-show'}
pubmed_prisma <- glue::glue(.open="<",.close=">",'subgraph cluster_1 {
    label="B. PubMed BAWTO Screen"
    style=filled;
    fillcolor=lightgrey;
    margin=8
    p_id    [label="Records identified\\non pubmed.gov\\n(n=<prisma$pm>)"];
    c2_id    [label="Records identified\\nfrom clinicaltrials.gov*\\n(n=<prisma$ctgov_ass>)", fillcolor=darkslategray1, style=filled];

    dedup     [label="Records after\\nduplicates removed\\n(n=<prisma$pm_screen>)",width=3]
    screening [label="Records screened\\n(n=<prisma$pm_screen>)",width=2]
    excluded  [label="Records excluded\\n(n=<prisma$pm_excl>)",width=2]

    c2_id -> dedup;
    p_id  -> dedup;
    dedup -> screening;
    screening -> excluded;

    fulltext_scr  [label="full-text articles assessed\\n for eligibility\\n(n=<prisma$fscreen>)"];
    fulltext_excl [label="Full-text articles excluded\\n (n=<prisma$fexcl>)\\n\\nno data for prostate cancer patients (n=<prisma$no_data_for_prostate_cancer_patients>)\\nno mutation data (n = <prisma$no_mutation_data>)\\nno outcome data (n=<prisma$no_outcome_data>)"];
    screening -> fulltext_scr;
    fulltext_scr -> fulltext_excl;

    fulltext_incl [label="studies included in\\nquantitative synthesis\\n(n=<prisma$synthesis>)", fillcolor=lightpink,style=filled];
    fulltext_scr -> fulltext_incl;
  }')
graph_prisma(pubmed_prisma)
```

## Final prisma diagram code:
The `prisma_figure` variable includes a few `{rank=same}` calls to move blocks around
```{r class.source = 'fold-show'}
final_prisma <- {
  glue::glue(.open="<",.close=">",
  'subgraph cluster_2 {
      label="C. HRD and DRD in mCRPC"
      margin=8

      other_records [label="studies in conference abstracts\\n and other reviews\\n(n=<prisma$conference>)"];
      fulltext_incl2 [label="studies included in\\nquantitative synthesis\\n(n=<prisma$synthesis>)", fillcolor=lightpink,style=filled];

      dedup_2 [label = "Records after\\nduplicates removed\\n(n=<prisma$synthesis2>)",width=3]
      hrd_screen [label="Records Screened\\n(n=<prisma$synthesis2>)"];
      hrd_excl [label="Records Excluded\\n(n=<prisma$synthesis2-prisma$trial.articles>)\\n\\nNo HRD/DRD biomarkers\\n(n=<prisma$synthesis2-prisma$trial.articles>)"];
      hrd_incl [label="studies included in\\nquantitative synthesis\\n(n=<prisma$trial.articles>)", fillcolor=lightgoldenrodyellow,style=filled];

      fulltext_incl2 -> dedup_2;
      other_records -> dedup_2;
      dedup_2 -> hrd_screen;
      hrd_screen -> hrd_excl;
      hrd_screen -> hrd_incl;

      { rank=same; hrd_screen; hrd_excl; }
    }'
)}

prisma_figure <- {
  glue::glue(.open="<",.close=">",
  'digraph prisma {
    newrank=true;
    node [shape="box", fontsize = 12, width=2];
    graph [compound=true,splines=ortho, nodesep=0.25, rankdir=TB];

    <clinicaltrials_prisma>
    <pubmed_prisma>
    <final_prisma>


    { rank=same; c_scr; c_excl; }
    { rank=same; p_id; c2_id;}
    { rank=same; screening; excluded;}
    { rank=same; fulltext_scr; fulltext_excl; }

  }')
}

DiagrammeR::grViz(prisma_figure)
```